Machine Learning Final Project
========================================================
by CCozzetto 

"Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways."


Loading Packages & Organizing Data
------------------------------------
```{r,message=FALSE, warning=FALSE}

install.packages("rpart.plot")
library(caret); library(rpart); library(rattle);library(randomForest);library(rpart.plot);library(RColorBrewer)

##Reading in downloaded data
testing <-read.csv("/Users/cynthiacozzetto/pml-testing.csv")
training <-read.csv("/Users/cynthiacozzetto/pml-training.csv")
summary(training$classe)

##Partitioning training & testing data from training set
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
training.1 <- training[inTrain, ]
testing.1 <- training[-inTrain, ]
set.seed(675)

##Clean Dataset for Descriptive Vars & N/As
length(colnames(training.1))
training.2 <- training.1[,colSums(is.na(training.1))<nrow(training.1)]
testing.2 <- training.1[,colSums(is.na(testing.1))<nrow(testing.1)]
length(colnames(training.2))

##Remove non-predictive colnames
training.3<-subset(training.2,select=7:length(colnames(training.2)))
testing.3<-subset(testing.2,select=7:length(colnames(testing.2)))

##Removing columns with NA values
na.table = sapply(training.3, function(x) {sum(is.na(x))})
table(na.table)
remove = names(na.table[na.table>13447])
training.4 = training.3[, !names(training.3) %in% remove]
testing.4 = training.3[, !names(testing.3) %in% remove]
```

There are 5 factor classes to be predicted.  The data set includes over 160 variables with some of the variables describing the data set (and not to be used as predictor variables).

Utilizing RPart & Random Forest to Build Model
----------------------------------------------
```{r}
##Building RPart model with selected variables from training.4
mod.rpart <- rpart(classe~pitch_arm + yaw_arm + roll_arm + roll_belt + pitch_belt + yaw_belt + gyros_belt_x + gyros_belt_y + gyros_belt_z + accel_belt_x + accel_belt_y + accel_belt_z + magnet_belt_x + magnet_belt_y + magnet_belt_z + gyros_arm_x + gyros_arm_y + gyros_arm_z + accel_arm_x + accel_arm_y + accel_arm_z + magnet_arm_x + magnet_arm_y + magnet_arm_z + roll_dumbbell + pitch_dumbbell + yaw_dumbbell, data=training.4,method="class")
##Visualizing Segmentation
fancyRpartPlot(mod.rpart)

pred.part <- predict(mod.rpart, testing.4, type = "class")
confusionMatrix(pred.part, training.4$classe)


##Random Forest Model
mod.rf <-randomForest(classe ~pitch_arm + yaw_arm + roll_arm + roll_belt + pitch_belt + yaw_belt + gyros_belt_x + gyros_belt_y + gyros_belt_z + accel_belt_x + accel_belt_y + accel_belt_z + magnet_belt_x + magnet_belt_y + magnet_belt_z + gyros_arm_x + gyros_arm_y + gyros_arm_z + accel_arm_x + accel_arm_y + accel_arm_z + magnet_arm_x + magnet_arm_y + magnet_arm_z + roll_dumbbell + pitch_dumbbell + yaw_dumbbell , data=training.4)
pred.rf <- predict(mod.rf, testing.4, type = "class")

##Confusion Matrix
confusionMatrix(pred.rf, testing.4$classe)

predictions <- predict(mod.rf, testing.4, type = "class")
predictions <- predict(mod.rf, testing, type = "class")
```

The confusion matrix for Decision Tree revels a low accuracy for predictions.

The confusion matrix for Random Forest reveals a very high accuracy for predicitions.
The 95% CI of accuracy between [.9997,1] results in an out of sample error of 0%.  This may indicate overfitting to the sample.

Submitting Assignment
-----------------------
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predictions)

##Code borrowed from instructions on project submission page.
```

